A survey on NoSQL and NewSQL datastores

### 2014-12-13 paper_dec/NoSQL_table.tex
moved table up a bit. Handing in.

\multicolumn{1}{|l|}{\textbf{Query optimization}}        & Query optimizer + Shard-keys to speed up distributed queries         & memcached                                                                                         & Bloom filter                                                                                     & Bloom filter                                                                                                                                      & Partition pruning using partition key & Queries in stored procedures planned at compile time                                                                           \\ \midrule

>>>>

\multicolumn{1}{|l|}{\textbf{Query optimization}}        & Query optimizer, shard-keys to speed up distributed queries         & memcached                                                                                         & Bloom filter                                                                                     & Bloom filter                                                                                                                                      & Partition pruning using partition key & Queries in stored procedures planned at compile time                                                                           \\ \midrule

### 2014-12-13 paper_dec/nosql_newsql_survey.tex
moved table up a bit. Handing in.

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. Because in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of generated data increases rapidly. For instance, the whole genome sequencing pipeline of the Broad Institute\cite{broad_institute}, a reference in the field, generates in the order of 3 TB of intermediary data when sequencing a single human genome. Also, when sequenced with 50x mean coverage (denoting the average number of times a base has been read during the process\cite{coverage_definition}) a single human genome takes 50 GB to store in compressed format\footnote{This information was gathered through personal interaction with researchers at the Exascience Life Lab}, so as this scales to genomes of millions of people and other organisms, storage and processing requirements become ever more demanding in terms of scalability, latency and concurrency.

@@

\item Key-value stores: much like dictionaries and associative maps, these map unique keys to values. Values are uninterpreted byte arrays, and the only way to access is by their key. This means modeling relations and complex structures between data, rich querying and indexing is not possible\cite{hecht2011nosql}\cite{grolinger2013data}. 

@@

\begin{landscape}
\includepdf[angle=90]{NoSQL_table.pdf}
\end{landscape}

@@

Because of the enormous choice of NoSQL and NewSQL datastores, an exhaustive study wasn't feasible. This survey reviews the most popular datastores in a few relevant categories, namely document stores, wide columnar stores and NewSQL stores. Key-value stores and graph databases were not taken into consideration, as their respective data models are not suited for the application at hand. A selection was then made based on similar criteria as in \cite{grolinger2013data}, following the ranking of DB-Engine Ranking \cite{db_engine_rank} as an indicator of popularity. This ranking tries to measure popularity based on a few parameters, such as the number of mentions on Web sites, general interest according to Google Trends, frequency of technical discussions on the Web, number of job listings, and number of professional profiles in which the systems are mentioned. The resulting selection consists of the document stores MongoDB and CouchBase Server, wide columnar stores Cassandra and HBase and NewSQL database VoltDB. There already exists an extension of the DNA sequencing pipeline used in the ExaScience Life Lab that allows for using MongoDB databases as input and/or output targets, making MongoDB even more relevant\cite{elprep_mongo}. Lastly, NewSQL query engine Cloudera Impala was also taken into consideration because of explicit interest from researchers in the aforementioned lab.
These 6 systems were then compared on characteristics relevant to high performance computing, such as indexing mechanisms, client interfaces to the data, distribution strategy, concurrency control and consistency models.

>>>>

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. Because in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of generated data increases rapidly. For instance, the whole genome sequencing pipeline of the Broad Institute\cite{broad_institute}, a reference in the field, generates in the order of 3 TB of intermediary data when sequencing a single human genome. Also, when sequenced with 50x mean coverage (denoting the average number of times a base has been read during the process\cite{coverage_definition}) a single human genome takes 50 GB to store in compressed format\footnote{This information was gathered during conversations with researchers at the Exascience Life Lab}, so as this scales to genomes of millions of people and other organisms, storage and processing requirements become ever more demanding in terms of scalability, latency and concurrency.

@@

\item Key-value stores: much like dictionaries and associative maps, these map unique keys to values. Values are uninterpreted byte arrays, and the only way to access them is by their key. This means modeling relations and complex structures between data, rich querying and indexing is not possible\cite{hecht2011nosql}\cite{grolinger2013data}. 

@@

@@

Because of the enormous choice of NoSQL and NewSQL datastores, an exhaustive study wasn't feasible. This survey reviews the most popular datastores in a few relevant categories, namely document stores, wide columnar stores and NewSQL stores. Key-value stores and graph databases were not taken into consideration, as their respective data models are not suited for the application at hand. A selection was then made based on similar criteria as in \cite{grolinger2013data}, following the ranking of DB-Engine Ranking \cite{db_engine_rank} as an indicator of popularity. 
\begin{landscape}
\includepdf[angle=90]{NoSQL_table.pdf}
\end{landscape}
This ranking tries to measure popularity based on a few parameters, such as the number of mentions on Web sites, general interest according to Google Trends, frequency of technical discussions on the Web, number of job listings, and number of professional profiles in which the systems are mentioned. The resulting selection consists of the document stores MongoDB and CouchBase Server, wide columnar stores Cassandra and HBase and NewSQL database VoltDB. There already exists an extension of the DNA sequencing pipeline used in the ExaScience Life Lab that allows for using MongoDB databases as input and/or output targets, making MongoDB even more relevant\cite{elprep_mongo}. Lastly, NewSQL query engine Cloudera Impala was also taken into consideration because of explicit interest from researchers in the aforementioned lab.
These 6 systems were then compared on characteristics relevant to high performance computing, such as indexing mechanisms, client interfaces to the data, distribution strategy, concurrency control and consistency models.
